---
title: "Lab 3/HW 4"
output: html_document
date: "2024-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Study Group
Chris and Isamari

## Tristate + BMI 

```{r}
library(ggplot2)
```


```{r}
library(tidyverse)
```
```{r}
brfss22$Age_midpt <- fct_recode(brfss22$X_AGEG5YR, "21" = "Age 18 to 24",
                                "27" = "Age 25 to 29", "32" = "Age 30 to 34",
                                "37" = "Age 35 to 39", "42" = "Age 40 to 44",
                                "47" = "Age 45 to 49", "52" = "Age 50 to 54",
                                "57" = "Age 55 to 59", "62" = "Age 60 to 64",
                                "67" = "Age 65 to 69", "72" = "Age 70 to 74",
                                "77" = "Age 75 to 79", "82" = "Age 80 or older",
                                NULL = "Dont know/Refused/Missing")

```

```{r}
brfss22$Age_midpt <- as.numeric(levels(brfss22$Age_midpt))[brfss22$Age_midpt]
```


```{r}
select_tristate <- (brfss22$X_STATE == "New York") | (brfss22$X_STATE == "New Jersey") | (brfss22$X_STATE == "Connecticut")
brfss_tristate <- subset(brfss22,select_tristate)
```

```{r}
choose_tristate <- (brfss22$X_STATE == "Alabama") | (brfss22$X_STATE == "Texas") | (brfss22$X_STATE == "Ohio")
brfss_tristate2 <- subset(brfss22,choose_tristate)
```


```{r}
p_tri <- ggplot(data = brfss_tristate,
                mapping = aes(x = Age_midpt,
                              y = X_BMI5))

```

```{r}
p_tri + geom_smooth()
```

```{r}
p_tri2 <- ggplot(data = brfss_tristate2,
                mapping = aes(x = Age_midpt,
                              y = X_BMI5))
```

```{r}
p_tri2 + geom_smooth()
```


```{r}
p_tri + geom_jitter(width = 2.5, height = NULL, alpha = 0.05)
```

```{r}
p_tri <- ggplot(data = brfss_tristate,
                mapping = aes(x = Age_midpt,
                              y = X_BMI5,
                              color = X_STATE,
                              fill = X_STATE))
```


```{r}
p_tri2 <- ggplot(data = brfss_tristate2,
                mapping = aes(x = Age_midpt,
                              y = X_BMI5,
                              color = X_STATE,
                              fill = X_STATE))
```

```{r}
p_tri2 + geom_smooth()
```



```{r}
p_tri + geom_smooth()
```


```{r}
p_tri3 <- ggplot(data = brfss_tristate,
                mapping = aes(x = Age_midpt,
                              y = X_BMI5,
                              color = X_STATE,
                              fill = X_STATE))
```

```{r}
p_tri3 + geom_point (color = "red") +
  geom_smooth(method = "loess") +
  scale_x_log10()
```

```{r}
psw <- ggplot(data = brfss_tristate,
              mapping = aes(x = SLEPTIM1,
                            y = Height_inches,
                            color = X_STATE,
                            fill = X_STATE))
```

```{r}
psw + geom_point() +
  geom_smooth(method = "loess") +
  scale_x_log10()
```



## Adverse Childhood Experiences


```{r}
ACEdidntask <- (as.numeric(is.na(brfss22$ACEDEPRS)) + 
                        as.numeric(is.na(brfss22$ACEDRINK)) +
                        as.numeric(is.na(brfss22$ACEDRUGS)) +
                        as.numeric(is.na(brfss22$ACEPRISN)) +
                        as.numeric(is.na(brfss22$ACEDIVRC)) +
                        as.numeric(is.na(brfss22$ACEPUNCH)) +
                        as.numeric(is.na(brfss22$ACEHURT1)) +
                        as.numeric(is.na(brfss22$ACESWEAR)) +
                        as.numeric(is.na(brfss22$ACETOUCH)) )
```



```{r}
select_ACE <- (ACEdidntask == 0)
```

```{r}
brfss_ACE <- subset(brfss22, select_ACE)

```


## ACE ALC AND DRUGS

```{r}
brfss_ACE$ACEDRINK_recode <- fct_recode(brfss_ACE$ACEDRINK, 
                                        "Y" = "Yes, Adverse Childhood Exper, lived with someone who was a problem drinker or alcoholic",
                                "N" = "No", 
                                NULL = "dont know not sure",
                                NULL = "refused"
                            
)

summary(brfss_ACE$ACEDRINK_recode)

```



```{r}
brfss_ACE$ACEDRUGS_recode <- fct_recode(brfss_ACE$ACEDRUGS, 
                                        "Y" = "Yes, Adverse Childhood Exper, lived with someone who used illegal street drugs or who abused prescription medications",
                                "N" = "No", 
                                NULL = "dont know not sure",
                                NULL = "refused"
                            
)

summary(brfss_ACE$ACEDRUGS_recode)
```



```{r}
xtabs(~ brfss_ACE$ACEDRINK_recode + brfss_ACE$ACEDRUGS_recode)
```


People who experienced ACE with alcohol and drugs - 6.71%
People who experienced ACE with only alcohol and no drugs - 17.30%
People who experienced ACE with only drugs and no alcohol - 3.14%
People who experienced neither - 71.94%


```{r}
ACEdidntask <- (as.numeric(is.na(brfss22$ACEDEPRS)) + 
                        as.numeric(is.na(brfss22$ACEDRINK)) +
                        as.numeric(is.na(brfss22$ACEDRUGS)) +
                        as.numeric(is.na(brfss22$ACEPRISN)) +
                        as.numeric(is.na(brfss22$ACEDIVRC)) +
                        as.numeric(is.na(brfss22$ACEPUNCH)) +
                        as.numeric(is.na(brfss22$ACEHURT1)) +
                        as.numeric(is.na(brfss22$ACESWEAR)) +
                        as.numeric(is.na(brfss22$ACETOUCH)) )
select_ACE <- (ACEdidntask == 0) # with zero missing values for any of ACE questions
```


## ACE DEPRESSION AND DRUGS

```{r}
levels(brfss_ACE$ACEDEPRS)
```



```{r}
brfss_ACE$ACEDEPRS_recode <- fct_recode(brfss_ACE$ACEDEPRS, 
                                        "Y" = "Yes, Adverse Childhood Exper, lived with someone who was depressed, mentally ill, or suicidal",
                                "N" = "No", 
                                NULL = "dont know not sure",
                                NULL = "refused"
                            
)

summary(brfss_ACE$ACEDEPRS_recode)
```

```{r}
xtabs(~ brfss_ACE$ACEDEPRS_recode + brfss_ACE$ACEDRUGS_recode)
```
People who had ACE with drugs and depressed family member - 5.97%
People who had ACE with drugs but no depressed family member - 3.88%
People who had ACE with depressed family member but no drugs - 12.56%
People who had neither ACE - 77.13%


## ACE DIVORCE AND DEPRESSION

```{r}
levels(brfss_ACE$ACEDIVRC)
```

```{r}
brfss_ACE$ACEDIVRC_recode <- fct_recode(brfss_ACE$ACEDIVRC, 
                            "Y" = "Yes, Adverse Childhood Exper, parents separated or divorced",
                                "N" = "No", 
                                NULL = "dont know not sure",
                                NULL = "refused", 
                                NULL = "parents never married"
                            
)

summary(brfss_ACE$ACEDIVRC_recode)
```

```{r}
xtabs(~ brfss_ACE$ACEDEPRS_recode + brfss_ACE$ACEDIVRC_recode)
```
People who had ACE with divorce/separation or family member with depression - 8.72%
People who had ACE with divorce/separation but no depressed family member - 17.17%
People who had ACE with depressed family member but no separation/divorce - 9.83%
People who had neither - 64.27%

```{r}

summary(brfss_ACE)
```


